---
title: "Scientific Journal and Sci-Fi Themed Color Palettes for ggplot2"
author: "Nan Xiao <<http://nanx.me>> <br>
         Miaozhu Li <<http://miaozhu.li>>"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
    css: ggsci.css
    includes:
      in_header: header.html
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Scientific Journal and Sci-Fi Themed Color Palettes for ggplot2}
---

# Introduction

> My eyes were finally opened and I understood nature.
>
> I learned at the same time to love it.
>
> --- Claude Monet

The `ggsci` package offers a collection of high-quality color palettes
inspired by colors used in scientific journals and science fiction TV shows.

# Available Palettes

The color palettes in `ggsci` are available as `ggplot2` scales.
For discrete palettes, the scales are named as:

  * `scale_color_palname()`
  * `scale_fill_palname()`

We also provided aliases, such as `scale_colour_palname()` for `scale_color_palname()`.
All available palettes are listed in the table below:

+----------------+------------------------------+-----------------+----------------------+
| Name           | Scales                       | Palette Types   | Palette Generator    |
+================+==============================+=================+======================+
| NPG            | `scale_color_npg()`          | `"nrc"`         | `pal_npg()`          |
|                | `scale_fill_npg()`           |                 |                      |
+----------------+------------------------------+-----------------+----------------------+
| AAAS           | `scale_color_aaas()`         | `"default"`     | `pal_aaas()`         |
|                | `scale_fill_aaas()`          |                 |                      |
+----------------+------------------------------+-----------------+----------------------+
| Lancet         | `scale_color_lancet()`       | `"lanonc"`      | `pal_lancet()`       |
|                | `scale_fill_lancet()`        |                 |                      |
+----------------+------------------------------+-----------------+----------------------+
| JCO            | `scale_color_jco()`          | `"default"`     | `pal_jco()`          |
|                | `scale_fill_jco()`           |                 |                      |
+----------------+------------------------------+-----------------+----------------------+
| UCSCGB         | `scale_color_ucscgb()`       | `"default"`     | `pal_ucscgb()`       |
|                | `scale_fill_ucscgb()`        |                 |                      |
+----------------+------------------------------+-----------------+----------------------+
| UChicago       | `scale_color_uchicago()`     | `"default"`     | `pal_uchicago()`     |
|                | `scale_fill_uchicago()`      | `"light"`       |                      |
|                |                              | `"dark"`        |                      |
+----------------+------------------------------+-----------------+----------------------+
| The Simpsons   | `scale_color_simpsons()`     | `"springfield"` | `pal_simpsons()`     |
|                | `scale_fill_simpsons()`      |                 |                      |
+----------------+------------------------------+-----------------+----------------------+
| Rick and Morty | `scale_color_rickandmorty()` | `"schwifty"`    | `pal_rickandmorty()` |
|                | `scale_fill_rickandmorty()`  |                 |                      |
+----------------+------------------------------+-----------------+----------------------+

Next, we use scatterplots with smooth curves, and bar plots to demonstrate the
discrete color palettes:

```{r}
library("ggsci")
library("ggplot2")
library("gridExtra")

data("diamonds")

p1 = ggplot(subset(diamonds, carat >= 2.2),
       aes(x = table, y = price, colour = cut)) +
  geom_point(alpha = 0.7) +
  geom_smooth(alpha = 0.05, size = 1, span = 1) +
  theme_bw()

p2 = ggplot(subset(diamonds, carat > 2.2 & depth > 55 & depth < 70), 
       aes(x = depth, fill = cut)) +
  geom_histogram(colour = "black", binwidth = 1, position = "dodge") +
  theme_bw()
```

## NPG

The NPG palette is inspired by journals published by Nature Publishing Group:

```{r, fig.width = 10.67, fig.height = 4, out.width = 800, out.height = 300, dpi = 150}
p1_npg = p1 + scale_color_npg()
p2_npg = p2 + scale_fill_npg()
grid.arrange(p1_npg, p2_npg, ncol = 2)
```

## AAAS

The AAAS palette is inspired by journals published by American Association
for the Advancement of Science:

```{r, fig.width = 10.67, fig.height = 4, out.width = 800, out.height = 300, dpi = 150}
p1_aaas = p1 + scale_color_aaas()
p2_aaas = p2 + scale_fill_aaas()
grid.arrange(p1_aaas, p2_aaas, ncol = 2)
```

## Lancet

The Lancet palette is inspired by Lancet journals, such as <emph>Lancet Oncology</emph>:

```{r, fig.width = 10.67, fig.height = 4, out.width = 800, out.height = 300, dpi = 150}
p1_lancet = p1 + scale_color_lancet()
p2_lancet = p2 + scale_fill_lancet()
grid.arrange(p1_lancet, p2_lancet, ncol = 2)
```

## JCO

The JCO palette is inspired by the colors used in <emph>Journal of Clinical Oncology</emph>:

```{r, fig.width = 10.67, fig.height = 4, out.width = 800, out.height = 300, dpi = 150}
p1_jco = p1 + scale_color_jco()
p2_jco = p2 + scale_fill_jco()
grid.arrange(p1_jco, p2_jco, ncol = 2)
```

## UCSCGB

The UCSCGB palette is from the colors used by
[UCSC Genome Browser](https://genome.ucsc.edu) for
representing chromosomes. This palette has been intensively
used in visualizations produced by [Circos](http://circos.ca).

```{r, fig.width = 10.67, fig.height = 4, out.width = 800, out.height = 300, dpi = 150}
p1_ucscgb = p1 + scale_color_ucscgb()
p2_ucscgb = p2 + scale_fill_ucscgb()
grid.arrange(p1_ucscgb, p2_ucscgb, ncol = 2)
```

## UChicago

The UChicago palette is based on
[the colors](http://communications.uchicago.edu/identity/color-palette)
used by The University of Chicago.
There are three palette types (`default`, `light`, `dark`) available.

```{r, fig.width = 10.67, fig.height = 12, out.width = 800, out.height = 900, dpi = 150}
p1_uchicago = p1 + scale_color_uchicago()
p2_uchicago = p2 + scale_fill_uchicago()
p1_uchicago_light = p1 + scale_color_uchicago("light")
p2_uchicago_light = p2 + scale_fill_uchicago("light")
p1_uchicago_dark  = p1 + scale_color_uchicago("dark")
p2_uchicago_dark  = p2 + scale_fill_uchicago("dark")
grid.arrange(p1_uchicago,       p2_uchicago,
             p1_uchicago_light, p2_uchicago_light,
             p1_uchicago_dark,  p2_uchicago_dark,
             ncol = 2, nrow = 3)
```

## The Simpsons

This palette is inspired by the colors used in the TV show <emph>The Simpsons</emph>:

```{r, fig.width = 10.67, fig.height = 4, out.width = 800, out.height = 300, dpi = 150}
p1_simpsons = p1 + scale_color_simpsons()
p2_simpsons = p2 + scale_fill_simpsons()
grid.arrange(p1_simpsons, p2_simpsons, ncol = 2)
```

## Rick and Morty

This palette is inspired by the colors used in the TV show <emph>Rick and Morty</emph>:

```{r, fig.width = 10.67, fig.height = 4, out.width = 800, out.height = 300, dpi = 150}
p1_rickandmorty = p1 + scale_color_rickandmorty()
p2_rickandmorty = p2 + scale_fill_rickandmorty()
grid.arrange(p1_rickandmorty, p2_rickandmorty, ncol = 2)
```

# Non-ggplot2 Graphics

To apply the color palettes in `ggsci` to other graphics systems
(such as base graphics and lattice graphics), simply use the
palette generator functions in the table above. For example:

```{r, fig.width = 6.67, fig.height = 6.67, out.width = 500, out.height = 500, dpi = 150}
mypal = pal_npg("nrc", alpha = 0.7)(9)
mypal

library("scales")
show_col(mypal)
```

You will be able to use the generated hex color codes for such
graphics systems accordingly. The transparent level of the
entire palette is easily adjustable via the argument `"alpha"`
in every generator or scale function.

# Discussion

Please note some of the palettes might not be the best choice for certain
purposes, such as color-blind safe, photocopy safe, or print friendly.
If you do have such considerations, you might want to check out
color palettes like [ColorBrewer](http://colorbrewer2.org)
and [viridis](https://cran.r-project.org/package=viridis).

The color palettes in this package are solely created for research purposes.
The authors are not responsible for the usage of such palettes.

<hr>

Project website: [ggsci.net](http://ggsci.net)
