---
title: "Scientific Journal and Sci-Fi Themed Color Palettes for ggplot2"
author: "Nan Xiao <<http://nanx.me>> <br>
         Miaozhu Li <<http://miaozhu.li>>"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
    css: ggsci.css
    includes:
      in_header: header.html
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Scientific Journal and Sci-Fi Themed Color Palettes for ggplot2}
---

# Introduction

> My eyes were finally opened and I understood nature.
>
> I learned at the same time to love it.
>
> --- Claude Monet

`ggsci` offers a collection of high-quality color palettes inspired by
colors used in scientific journals, data visualization libraries,
and science fiction TV shows. The color palettes in `ggsci` are
available as `ggplot2` scales. For all the color palettes,
the corresponding scales are named as:

  * `scale_color_palname()`
  * `scale_fill_palname()`

We also provided aliases, such as `scale_colour_palname()` for
`scale_color_palname()`. All available color palettes are
summarized in the table below.

+----------------+------------------------------+-------------------+----------------------+
| Name           | Scales                       | Palette Types     | Palette Generator    |
+================+==============================+===================+======================+
| NPG            | `scale_color_npg()`          | `"nrc"`           | `pal_npg()`          |
|                | `scale_fill_npg()`           |                   |                      |
+----------------+------------------------------+-------------------+----------------------+
| AAAS           | `scale_color_aaas()`         | `"default"`       | `pal_aaas()`         |
|                | `scale_fill_aaas()`          |                   |                      |
+----------------+------------------------------+-------------------+----------------------+
| Lancet         | `scale_color_lancet()`       | `"lanonc"`        | `pal_lancet()`       |
|                | `scale_fill_lancet()`        |                   |                      |
+----------------+------------------------------+-------------------+----------------------+
| JCO            | `scale_color_jco()`          | `"default"`       | `pal_jco()`          |
|                | `scale_fill_jco()`           |                   |                      |
+----------------+------------------------------+-------------------+----------------------+
| UCSCGB         | `scale_color_ucscgb()`       | `"default"`       | `pal_ucscgb()`       |
|                | `scale_fill_ucscgb()`        |                   |                      |
+----------------+------------------------------+-------------------+----------------------+
| D3             | `scale_color_d3()`<br>       | `"category10"`    | `pal_d3()`           |
|                | `scale_fill_d3()`            | `"category20"`    |                      |
|                |                              | `"category20b"`   |                      |
|                |                              | `"category20c"`   |                      |
+----------------+------------------------------+-------------------+----------------------+
| UChicago       | `scale_color_uchicago()`     | `"default"`<br>   | `pal_uchicago()`     |
|                | `scale_fill_uchicago()`      | `"light"`<br>     |                      |
|                |                              | `"dark"`          |                      |
+----------------+------------------------------+-------------------+----------------------+
| The Simpsons   | `scale_color_simpsons()`     | `"springfield"`   | `pal_simpsons()`     |
|                | `scale_fill_simpsons()`      |                   |                      |
+----------------+------------------------------+-------------------+----------------------+
| Futurama       | `scale_color_futurama()`     | `"planetexpress"` | `pal_futurama()`     |
|                | `scale_fill_futurama()`      |                   |                      |
+----------------+------------------------------+-------------------+----------------------+
| Rick and Morty | `scale_color_rickandmorty()` | `"schwifty"`      | `pal_rickandmorty()` |
|                | `scale_fill_rickandmorty()`  |                   |                      |
+----------------+------------------------------+-------------------+----------------------+
| GSEA           | `scale_color_gsea()`         | `"default"`       | `pal_gsea()`         |
|                | `scale_fill_gsea()`          |                   |                      |
+----------------+------------------------------+-------------------+----------------------+

# Discrete Color Palettes

We will use scatterplots with smooth curves, and bar plots to demonstrate
the discrete color palettes in `ggsci`.

```{r}
library("ggsci")
library("ggplot2")
library("gridExtra")

data("diamonds")

p1 = ggplot(subset(diamonds, carat >= 2.2),
       aes(x = table, y = price, colour = cut)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "loess", alpha = 0.05, size = 1, span = 1) +
  theme_bw()

p2 = ggplot(subset(diamonds, carat > 2.2 & depth > 55 & depth < 70), 
       aes(x = depth, fill = cut)) +
  geom_histogram(colour = "black", binwidth = 1, position = "dodge") +
  theme_bw()
```

## NPG

The NPG palette is inspired by the plots in journals published by Nature Publishing Group:

```{r, fig.width = 10.67, fig.height = 4, out.width = 800, out.height = 300, dpi = 150}
p1_npg = p1 + scale_color_npg()
p2_npg = p2 + scale_fill_npg()
grid.arrange(p1_npg, p2_npg, ncol = 2)
```

## AAAS

The AAAS palette is inspired by the plots in journals published by American Association
for the Advancement of Science:

```{r, fig.width = 10.67, fig.height = 4, out.width = 800, out.height = 300, dpi = 150}
p1_aaas = p1 + scale_color_aaas()
p2_aaas = p2 + scale_fill_aaas()
grid.arrange(p1_aaas, p2_aaas, ncol = 2)
```

## Lancet

The Lancet palette is inspired by the plots in Lancet journals, such as <emph>Lancet Oncology</emph>:

```{r, fig.width = 10.67, fig.height = 4, out.width = 800, out.height = 300, dpi = 150}
p1_lancet = p1 + scale_color_lancet()
p2_lancet = p2 + scale_fill_lancet()
grid.arrange(p1_lancet, p2_lancet, ncol = 2)
```

## JCO

The JCO palette is inspired by the the plots in <emph>Journal of Clinical Oncology</emph>:

```{r, fig.width = 10.67, fig.height = 4, out.width = 800, out.height = 300, dpi = 150}
p1_jco = p1 + scale_color_jco()
p2_jco = p2 + scale_fill_jco()
grid.arrange(p1_jco, p2_jco, ncol = 2)
```

## UCSCGB

The UCSCGB palette is from the colors used by
[UCSC Genome Browser](https://genome.ucsc.edu) for
representing chromosomes. This palette has been intensively
used in visualizations produced by [Circos](http://circos.ca).

```{r, fig.width = 10.67, fig.height = 4, out.width = 800, out.height = 300, dpi = 150}
p1_ucscgb = p1 + scale_color_ucscgb()
p2_ucscgb = p2 + scale_fill_ucscgb()
grid.arrange(p1_ucscgb, p2_ucscgb, ncol = 2)
```

## D3

The D3 palette is from the categorical colors used by
[D3.js](https://d3js.org) (version 3.x and before).
There are four palette types (`category10`, `category20`,
`category20b`, `category20c`) available.

```{r, fig.width = 10.67, fig.height = 16, out.width = 800, out.height = 1200, dpi = 150}
p1_d3 = p1 + scale_color_d3()
p2_d3 = p2 + scale_fill_d3()
p1_d3_c10 = p1 + scale_color_d3("category10")
p2_d3_c10 = p2 + scale_fill_d3("category10")
p1_d3_c20 = p1 + scale_color_d3("category20")
p2_d3_c20 = p2 + scale_fill_d3("category20")
p1_d3_c20b = p1 + scale_color_d3("category20b")
p2_d3_c20b = p2 + scale_fill_d3("category20b")
p1_d3_c20c = p1 + scale_color_d3("category20c")
p2_d3_c20c = p2 + scale_fill_d3("category20c")
grid.arrange(p1_d3_c10,  p2_d3_c10,
             p1_d3_c20,  p2_d3_c20,
             p1_d3_c20b, p2_d3_c20b,
             p1_d3_c20c, p2_d3_c20c,
             ncol = 2, nrow = 4)
```

## UChicago

The UChicago palette is based on
[the colors](http://communications.uchicago.edu/identity/color-palette)
used by The University of Chicago.
There are three palette types (`default`, `light`, `dark`) available.

```{r, fig.width = 10.67, fig.height = 12, out.width = 800, out.height = 900, dpi = 150}
p1_uchicago = p1 + scale_color_uchicago()
p2_uchicago = p2 + scale_fill_uchicago()
p1_uchicago_light = p1 + scale_color_uchicago("light")
p2_uchicago_light = p2 + scale_fill_uchicago("light")
p1_uchicago_dark  = p1 + scale_color_uchicago("dark")
p2_uchicago_dark  = p2 + scale_fill_uchicago("dark")
grid.arrange(p1_uchicago,       p2_uchicago,
             p1_uchicago_light, p2_uchicago_light,
             p1_uchicago_dark,  p2_uchicago_dark,
             ncol = 2, nrow = 3)
```

## The Simpsons

This palette is inspired by the colors used in the TV show <emph>The Simpsons</emph>:

```{r, fig.width = 10.67, fig.height = 4, out.width = 800, out.height = 300, dpi = 150}
p1_simpsons = p1 + scale_color_simpsons()
p2_simpsons = p2 + scale_fill_simpsons()
grid.arrange(p1_simpsons, p2_simpsons, ncol = 2)
```

## Futurama

This palette is inspired by the colors used in the TV show <emph>Futurama</emph>:

```{r, fig.width = 10.67, fig.height = 4, out.width = 800, out.height = 300, dpi = 150}
p1_futurama = p1 + scale_color_futurama()
p2_futurama = p2 + scale_fill_futurama()
grid.arrange(p1_futurama, p2_futurama, ncol = 2)
```

## Rick and Morty

This palette is inspired by the colors used in the TV show <emph>Rick and Morty</emph>:

```{r, fig.width = 10.67, fig.height = 4, out.width = 800, out.height = 300, dpi = 150}
p1_rickandmorty = p1 + scale_color_rickandmorty()
p2_rickandmorty = p2 + scale_fill_rickandmorty()
grid.arrange(p1_rickandmorty, p2_rickandmorty, ncol = 2)
```

# Continuous Color Palettes

We will use a correlation matrix visualization (a special type of heatmap)
to demonstrate the continuous color palettes in `ggsci`.

```{r}
library("reshape2")

data("mtcars")
cor = cor(mtcars)
cor_melt = melt(cor)

p3 = ggplot(cor_melt,
            aes(x = Var1, y = Var2, fill = value)) + 
  geom_tile(colour = "black", size = 0.3) +
  theme_bw()
```

## GSEA

The GSEA palette (continuous) is inspired by the heatmaps generated by
[GSEA GenePattern](http://software.broadinstitute.org/cancer/software/genepattern/).

```{r, fig.width = 8, fig.height = 6.67, out.width = 600, out.height = 500, dpi = 150}
p3_gsea = p3 + scale_fill_gsea()
p3_gsea
```

# Non-ggplot2 Graphics

To apply the color palettes in `ggsci` to other graphics systems
(such as base graphics and lattice graphics), simply use the
palette generator functions in the table above. For example:

```{r, fig.width = 6.67, fig.height = 6.67, out.width = 500, out.height = 500, dpi = 150}
mypal = pal_npg("nrc", alpha = 0.7)(9)
mypal

library("scales")
show_col(mypal)
```

You will be able to use the generated hex color codes for such
graphics systems accordingly. The transparent level of the
entire palette is easily adjustable via the argument `"alpha"`
in every generator or scale function.

# Discussion

Please note some of the palettes might not be the best choice for certain
purposes, such as color-blind safe, photocopy safe, or print friendly.
If you do have such considerations, you might want to check out
color palettes like [ColorBrewer](http://colorbrewer2.org)
and [viridis](https://cran.r-project.org/package=viridis).

The color palettes in this package are solely created for research purposes.
The authors are not responsible for the usage of such palettes.

<hr>

Project website: [ggsci.net](http://ggsci.net)
